---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Data visualisation: `ggplot2` {#ggplot2}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library("tidyverse")
```

## Why visualising data?
### The Anscombe’s Quartet
In Anscombe, F. J. (1973). "Graphs in Statistical Analysis" there was the following dataset:

```{r}
quartet <- read_csv("https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/anscombe.csv")
quartet
quartet %>% 
  group_by(dataset) %>% 
  summarise(mean_X = mean(x),
            mean_Y = mean(y),
            sd_X = sd(x),
            sd_Y = sd(y),
            cor = cor(x, y),
            n_obs = n()) %>% 
  select(-dataset) %>% 
  round(2)
```

Lets visualise those datasets:

```{r, echo = FALSE}
quartet %>% 
  ggplot(aes(x, y))+
  geom_point()+
  geom_smooth(method = "lm", se = F)+
  facet_wrap(~dataset)+
  theme_bw()
```

### The DataSaurus
In [Matejka and Fitzmaurice (2017) "Same Stats, Different Graphs"](https://www.autodeskresearch.com/sites/default/files/SameStats-DifferentGraphs.pdf) there are the following datasets:

```{r}
datasaurus <- read_csv("https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/datasaurus.csv")
datasaurus
```

```{r, echo=FALSE}
datasaurus %>% 
  ggplot(aes(x, y))+
  geom_point()+
  facet_wrap(~dataset)+
  theme_bw()
```

And... all discriptive statistics are the same!

```{r}
datasaurus %>% 
  group_by(dataset) %>% 
  summarise(mean_X = mean(x),
            mean_Y = mean(y),
            sd_X = sd(x),
            sd_Y = sd(y),
            cor = cor(x, y),
            n_obs = n()) %>% 
  select(-dataset) %>% 
  round(1)
```

## Basic `ggplot2`

`ggplot2` is a modern tool for data visualisation. There are [a lot of extentions](http://www.ggplot2-exts.org/gallery/) for `ggplot2`. There is also [a cheatsheet on `ggplot2`](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf). There is also a whole book about `ggplot2` [@wickham16].

Every `ggplot2` plot has three key components:

* data,
* A set of aesthetic mappings between variables in the data and visual properties, and
* At least one layer which describes how to render each observation. Layers
are usually created with a `geom_...()` function.

### Scaterplot

I downloaded a Polish dictionary from [here](https://sjp.pl/slownik/odmiany/). I removed all abbreviations and proper names and took only one form from the paradigm. After all this I calculated the number of syllables (simply by counting vowels, combinations of *i* and other vowels I counted as one), number of symbols in each word and extracted the first letter. Here is the [result dataset](https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/polish_dictionary.csv).

```{block, type = "rmdtask"}
Download this dataset to the variable `polish_dictionary`. How many words are there?
```

```{r, include=FALSE}
polish_dictionary <- read_csv("https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/polish_dictionary.csv")
```


```{r, results='asis', echo = FALSE, eval = knitr::is_html_output()}
checkdown::autocheck_question(question_id = "3_1", answer =  nrow(polish_dictionary))
```

So this data could be visualised using the following code:

* `ggplot2`
```{r, eval = FALSE}
ggplot(data = polish_dictionary, aes(x = n_char, y = n_vowels)) +
  geom_point()
```

* `dplyr` and `ggplot2`
```{r first_ggplot, cache=TRUE}
polish_dictionary %>%
  ggplot(aes(x = n_char, y = n_vowels))+
  geom_point()
```

#### Layers

All commands in `ggplot2` are separated by `+` sign (author of the package, Hadley Wickham, [is deeply regrets that it is not `%>%`](https://www.youtube.com/watch?v=vYwXMnC03I4)), but their order matters:

```{r layers_1, cache=TRUE}
polish_dictionary %>%
  ggplot(aes(n_char, n_vowels))+
  geom_point()+
  geom_smooth()
```

```{r layers_2, cache=TRUE}
polish_dictionary %>%
  ggplot(aes(n_char, n_vowels))+
  geom_smooth()+
  geom_point()
```

#### `aes()`

Since every `ggplot2` plot has data as a key components there is a function `aes()` that maps variables from dataframe onto visual properties of the graph. There is a simple rule:

**If values are from dataframe put them into `aes()`, otherwise --- don't.**

```{r aes_1, cache=TRUE}
polish_dictionary %>%
  ggplot(aes(n_char, n_vowels, color = first_letter))+
  geom_point()
```

```{r aes_2, cache=TRUE}
polish_dictionary %>%
  ggplot(aes(n_char, n_vowels))+
  geom_point(color = "green")
```

```{r, cache=TRUE}
polish_dictionary %>%
  ggplot(aes(n_char, n_vowels))+
  geom_point(aes(color = first_letter))
```

There are some other possibilities to mark categories:

* with `shape` argument

```{r shape, cache=TRUE}
polish_dictionary %>%
  filter(first_letter == "a" |
           first_letter == "i" |
           first_letter == "u") %>% 
  ggplot(aes(n_char, n_vowels, shape = first_letter))+
  geom_point()
```

* with `label` argument and `geom_text()`

```{r label, cache=TRUE}
polish_dictionary %>%
  filter(first_letter == "a" |
           first_letter == "i" |
           first_letter == "u") %>% 
  ggplot(aes(n_char, n_vowels, label = first_letter))+
  geom_text()
```

Sometimes annotations overlap:
```{r}
polish_dictionary %>%
  slice(8400:8450) %>%  # lets pick 50 words from our dictionary
  ggplot(aes(n_char, n_vowels, label = word))+
  geom_text()
```

Then it is better to use `geom_text_repel()` from the `ggrepel` library (do not forget to download it using `install.packages("ggrepel")`):
```{r}
library("ggrepel")
polish_dictionary %>%
  slice(8400:8450) %>% 
  ggplot(aes(n_char, n_vowels, label = word))+
  geom_text_repel()
```

It looks better, when you add some points:

```{r}
polish_dictionary %>%
  slice(8400:8450) %>% 
  ggplot(aes(n_char, n_vowels, label = word))+
  geom_text_repel()+
  geom_point()
```

#### Annotate labels, axis, caption etc.

```{r}
polish_dictionary %>%
  slice(8400:8450) %>% 
  ggplot(aes(n_char, n_vowels, label = word))+
  geom_text_repel()+
  geom_point()+
  labs(x = "number of characters",
       y = "number of vowels",
       title = "Correlation of number of characters and vowels",
       subtitle = "data from sjp.pl",
       caption = "this graph is created with ggplot2")
```

```{block, type = "rmdtask"}
Download [this dataset](https://raw.githubusercontent.com/agricolamz/DS_for_DH/master/data/first_scatterplot.csv) and create a scatterplot. What is there?
```

```{r, results='asis', echo = FALSE, eval = knitr::is_html_output()}
checkdown::autocheck_question(question_id = "3_2", answer =  "DH")
```

### Barplots

The same data can be aggregated and non-aggregated:

```{r}
misspelling <- read_csv("https://raw.githubusercontent.com/agricolamz/DS_for_DH/master/data/misspelling_dataset.csv")
misspelling
```

* variable `spelling` is **aggregated**: for each value of `speeling` variable there is a corresponging value in `count` variable.
* variable `correct` is **non-aggregated**: there is no any variable associated with counts of `correct` variable
