# Data visualisation: `ggplot2` {#ggplot2}

```{r, message = FALSE}
library("tidyverse")
```

## Why visualise data?
### The Anscombe’s Quartet
In Anscombe, F. J. (1973). "Graphs in Statistical Analysis" there were the following dataset:

```{r, message= FALSE}
quartet <- read_csv("https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/anscombe.csv")
quartet
quartet %>% 
  group_by(dataset) %>% 
  summarise(mean_X = mean(x),
            mean_Y = mean(y),
            sd_X = sd(x),
            sd_Y = sd(y),
            cor = cor(x, y),
            n_obs = n()) %>% 
  select(-dataset) %>% 
  round(2)
```

Lets visualise those datasets:

```{r, echo = FALSE, message=FALSE}
quartet %>% 
  ggplot(aes(x, y))+
  geom_point()+
  geom_smooth(method = "lm", se = F)+
  facet_wrap(~dataset)+
  theme_bw()
```

### The DataSaurus
In [Matejka and Fitzmaurice (2017) "Same Stats, Different Graphs"](https://www.autodeskresearch.com/sites/default/files/SameStats-DifferentGraphs.pdf) there are the following datasets:

```{r, message= FALSE}
datasaurus <- read_csv("https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/datasaurus.csv")
datasaurus
```

```{r, echo=FALSE, message= FALSE}
datasaurus %>% 
  ggplot(aes(x, y))+
  geom_point()+
  facet_wrap(~dataset)+
  theme_bw()
```

And... all discriptive statistics are the same!

```{r}
datasaurus %>% 
  group_by(dataset) %>% 
  summarise(mean_X = mean(x),
            mean_Y = mean(y),
            sd_X = sd(x),
            sd_Y = sd(y),
            cor = cor(x, y),
            n_obs = n()) %>% 
  select(-dataset) %>% 
  round(1)
```

## Basic `ggplot2`

`ggplot2` is a modern tool for data visualisation. There are [a lot of extentions](http://www.ggplot2-exts.org/gallery/) for `ggplot2`. There is also [a cheatsheet on `ggplot2`](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf). There is also a whole book about `ggplot2` [@wickham16].

Every `ggplot2` plot has three key components:

* data,
* A set of aesthetic mappings between variables in the data and visual properties, and
* At least one layer which describes how to render each observation. Layers
are usually created with a `geom_...()` function.

### Scaterplot

I downloaded a Polish dictionary from [here](https://sjp.pl/slownik/odmiany/). I removed all abbreviations and proper names and took only one form from the paradigm. After all this I calculated the number of syllables (simply by countyng vowels, combinations of *i* and other vowels I counted as one) and number of symbols in each word. Here is the [result dataset](https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/polish_dictionary.csv).

```{block, type = "rmdtask"}
Download this dataset to the variable `polish_dictionary`. How many words are there?
```

```{r, include=FALSE}
polish_dictionary <- read_csv("https://raw.githubusercontent.com/agricolamz/2020.02_Naumburg_R/master/data/polish_dictionary.csv")
```

So this data could be visualised using the following code:

* `ggplot2`
```{r, eval = FALSE}
ggplot(data = polish_dictionary, aes(n_char, n_vowels)) +
  geom_point()
```

* `dplyr`, `ggplot2`
```{r first_ggplot, cache=TRUE}
polish_dictionary %>%
  ggplot(aes(n_char, n_vowels))+
  geom_point()
```



